---
title: "Visual attention to neutral vs competing vs reinforcing stimuli among children with ASD"
output: html_notebook
---
<b>Author: Dominik Neumann</b><br>
Michigan State University

```{r global_options, include = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Setup
Please select "dat_visualattention.csv" for dat_attention data table and "dat_master.csv" for dat_master data table in order to run this script properly. <br>
Missing values were coded as -999. We coded “-“ values as -888, indicating participants looked at the center of the screen and thus at neither distracter nor target. We coded “.” values as -777, indicating participants looking outside of the screen.
```{r results = FALSE}
### OPTIONS
options(scipen = 3)

### LOADING PACKAGES
# General libraries
library(tidyverse)
library(psych)
library(plyr)

# Multilevel modeling
library(lme4)
library(nlme)
library(ez)
library(multcomp)
library(TTR)
library(lmerTest)
library(powerlmm)
library(lsmeans)
library(car)

# ANOVA
library(knitr)
library(sjstats)

### FUNCTIONS
# APA style charts
apatheme <- theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(),
        text = element_text(family = 'Times'))

# Function that counts the number if NA's in a given row:
countNA <- function(df)
  apply(df, MARGIN = 1, FUN = function(x) length(x[is.na(x)]))

### DATA
# Loading gaze data
dat_attention <- read.csv(file.choose())
dat_master <- read.csv(file.choose())

# Recoding '-999', '-888', and '-777' to missing values 'NA'
dat_attention[dat_attention == -999 | dat_attention == -888 | dat_attention == -777] <- NA
```