---
title: "Visual attention to neutral vs competing vs reinforcing stimuli among children with ASD"
output: html_notebook
---
<b>Author: Dominik Neumann</b><br>
Michigan State University

```{r global_options, include = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Setup

Please select "dat_visualattention.csv" for dat_attention data table and "dat_master.csv" for dat_master data table in order to run this script properly. <br>
Missing values were coded as -999. We coded “-“ values as -888, indicating participants looked at the center of the screen and thus at neither distracter nor target. We coded “.” values as -777, indicating participants looking outside of the screen.

```{r results = FALSE}
### OPTIONS
options(scipen = 3)

### LOADING PACKAGES
# General libraries
library(tidyverse)
library(psych)
library(plyr)

# Multilevel modeling
library(lme4)
library(nlme)
library(ez)
library(multcomp)
library(optimx)
library(TTR)
library(lmerTest)
library(powerlmm)
library(lsmeans)
library(car)

# ANOVA
library(knitr)
library(sjstats)

### FUNCTIONS
# APA style charts
apatheme <- theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(),
        text = element_text(family = 'Times'))

# Function that counts the number if NA's in a given row:
countNA <- function(df)
  apply(df, MARGIN = 1, FUN = function(x) length(x[is.na(x)]))

### DATA
# Loading gaze data
dat_attention <- read.csv(file.choose())
dat_master <- read.csv(file.choose())

# Recoding '-999', '-888', and '-777' to missing values 'NA'
dat_attention[dat_attention == -999 | dat_attention == -888 | dat_attention == -777] <- NA
```

There are 58 participants in this study that have been diagnosed with Autism Syndrom Disorder (ASD). All conditions were administered within participants, with ten trials per condition.

```{r results = FALSE}
# We were only interested in participants with with Autism Syndrom Disorder
dat_attention$Child_ID <- factor(dat_attention$Child_ID)
dat_attention <-dat_attention %>%
  dplyr::filter(Child_Type == "ASD")
nrow(count(unique(dat_attention$Child_ID)))
```

# Data cleaning and transforming
## Missing values and underperforming trials
Our critical window is from 300 to 2000 ms after the onset of the target word. Trials on which children were not fixating either object for more than 50% of the critical window were excluded. These trials do not have adequate data and at the extreme could include instances where children looked at one object for a single frame at the start of the window and then looked away (and would therefore have 0 or 100% accuracy).
```{r results = FALSE}
dat_attention_window <- subset(dat_attention, select = -c(frame_001:frame_096, frame_149:frame_200))

dat_attention_window_trialNA <- dat_attention_window[countNA(subset(dat_attention_window, select = frame_097:frame_148)) < (ncol(subset(dat_attention_window, select = frame_097:frame_148))/2),]
```

There were n = 397 trials with more than 50% missing values across variables assed, these were deleted, leading to 916 observed trials.
<b> Table 1 </b>
```{r}
Process <- c("Before cleaning", "Number of excluded trials", "After cleaning")
n <- c(nrow(dat_attention_window), nrow(dat_attention_window) - nrow(dat_attention_window_trialNA), nrow(dat_attention_window_trialNA))
number_of_observations <- cbind(Process, n)
as.data.frame(number_of_observations)
```

```{r results = FALSE}
`%notin%` <- Negate(`%in%`)

childxconditioncount_pre <- count(dat_attention_window_trialNA, c("Child_ID", "Condition"))
dat_attention_critical_trialNA <- dat_attention_window_trialNA %>%
  dplyr::filter(Child_ID %notin% c("8014", "8016", "8022", "8031", "8036", "8044", "8047", "8049", "8050", "8055", "8057", "8058", "8059", "8061",  "8062", "8063", "8072", "8074", "8076", "8077"))
childxconditioncount_post <- count(dat_attention_window_trialNA, c("Child_ID", "Condition"))
```

There were N = 56 participants who provided data. However, 20 participants contributed less than 4 trials per neutral/competing/reinforcing condition respectively. These were excluded in the next step. Thus, after excluding participants who contributed less than four trials to the experiment the new number of participants is N = 36. <p>
<p>
<b> Table 2 </b>
```{r}
Process <- c("Before cleaning", "After cleaning")
N <- c(length(unique(childxconditioncount_pre$Child_ID)), length(unique(childxconditioncount_post$Child_ID)))
number_of_participants <- cbind(Process, N)
as.data.frame(number_of_participants)
```

## Overview of missing data

A median of 3 values (M = 4.60, SD = 5.84) were missing per participant for the eye tracking data (participant did not look at either of the pictures). Table 3 summarizes overall sample and Table 4 summarizes missing values by condition. <br>

<b>Table 3 </b>
```{r results = FALSE}
dat_attention_window_descriptive <- dat_attention_window_trialNA %>%
  dplyr::select("Condition", starts_with("frame_"))
dat_attention_window_descriptive$NAcount <- countNA(dat_attention_window_descriptive)
describe(dat_attention_window_descriptive$NAcount)
```

<b>Table 4 </b>
```{r}
descriptive_missing <- describeBy(dat_attention_window_descriptive$NAcount,
                                  dat_attention_window_descriptive$Condition,
                                  mat = TRUE, digits = 2)
descriptive_missing
```

## Overview of trials per participant
A median of 7.5 trials were completed per participant overall (M = 7.40, SD = 1.84). Table 5 summarizes completed trials by condition. <br>

<b>Table 5 </b>
```{r results = FALSE}
dat_attention_window_trialNA$Child_Condition <- interaction(dat_attention_window_trialNA$Child_ID,
                                                            dat_attention_window_trialNA$Condition)

dat_counts <- ddply(dat_attention_window_trialNA, ~Child_Condition, summarise, Counts_TR = length(unique(Tr_Num)))
dat_attention_window_trialNA <- merge(dat_attention_window_trialNA, dat_counts, by = "Child_Condition")

descriptive_trial <- describeBy(dat_attention_window_trialNA$Counts_TR,
                                dat_attention_window_trialNA$Condition,
                                mat = TRUE, digits = 2)
descriptive_trial
```

## Transformation to long data

```{r results = FALSE}
# Aliasing data
dat_analysiswindow <- dat_attention_window_trialNA
```

In order to analyse data appropriately, data was transformed to a series of long datasets. First, a dataset for the analysis window only (frame 97 to frame 148) was created. Then, a dataset for all time points (frame 1 to frame 200) was created. Additionally, a dataset was generated that included the time before the analysis window as well as after the analysis window.<br>

```{r results = FALSE}
### TRANSFORMING ANALYSIS WINDOW ONLY
dat_analysiswindow$Child_ID <- factor(dat_analysiswindow$Child_ID)
dat_analysiswindow_long <- gather(dat_analysiswindow, frame, accuracy, frame_097:frame_148, factor_key = TRUE)

### OVERALL DATA
# Subsetting full data based on prior cleaning procedures
dat_attention_full <- subset(dat_attention, Child_ID %in% dat_analysiswindow$Child_ID)

# Subsetting for the pre-analysis window
dat_attention_pre <- subset(dat_attention_full, select = -c(frame_097:frame_200))
dat_attention_pre$prepost <- "pre-analysis"
dat_attention_pre$Child_ID <- factor(dat_attention_pre$Child_ID)
dat_attention_pre_long <- gather(dat_attention_pre, frame, accuracy, frame_001:frame_096, factor_key = TRUE)

# Subsetting for the analysis window
dat_attention_analysis <- subset(dat_attention_full, select = -c((frame_001:frame_096), (frame_149:frame_200)))
dat_attention_analysis$prepost <- "analysis"
dat_attention_analysis$Child_ID <- factor(dat_attention_analysis$Child_ID)
dat_attention_analysis_long <- gather(dat_attention_analysis, frame, accuracy, frame_097:frame_148, factor_key = TRUE)

# Subsetting for the post-analysis window
dat_attention_post <- subset(dat_attention_full, select = -c((frame_001:frame_148)))
dat_attention_post$prepost <- "post-analysis"
dat_attention_post$Child_ID <- factor(dat_attention_post$Child_ID)
dat_attention_post_long <- gather(dat_attention_post, frame, accuracy, frame_149:frame_200, factor_key = TRUE)

# Combining
dat_attention_long <- rbind(dat_attention_pre_long, dat_attention_analysis_long, dat_attention_post_long)
```

# Measuring Accuracy

"ACCURACY" is measured as '0', '1', or 'NA'.<br>
1 = participant focuses on target at specific frame<br>
0 = participant focuses on the distracter at specific frame<br>
NA = participant focuses either on a point outside the screen (wall, machines, roof, etc.) or on the center of the screen (neither picture) at specific frame<br>
"MEAN ACCURACY" is measured as the mean of accuracy observations across individuals for each frame within a given conditions.

```{r results = FALSE}
# For the analysis window dataframe
dat_analysiswindow_long$accuracy <- as.numeric(as.character(dat_analysiswindow_long$accuracy))
dat_analysiswindow_long <- dat_analysiswindow_long %>%
  dplyr::group_by(Child_ID, frame, Condition) %>%
  dplyr::mutate(mean_accuracy = mean(accuracy, na.rm = TRUE))
# For the overall dataframe
dat_attention_long$accuracy <- as.numeric(as.character(dat_attention_long$accuracy))
dat_attention_long <- dat_attention_long %>%
  dplyr::group_by(Child_ID, frame, Condition) %>%
  dplyr::mutate(mean_accuracy = mean(accuracy, na.rm = TRUE))
```

# Plotting Mean Accuracy by Condition

Figure 1 depicts development of mean accuracy. That is, the y-axis describes mean accuracy for all participants over all trials and the x-axis describes the time from beginning to end of the experiment. Mean accuracy measures are stratified by condition. We can see that children generally paid more attention to the correct stimulus in the "reinforcing condition". Children further paid the least attention to the correct stimulus in the competing condition. However, an interesting orienting pattern emerged in the beginning, where participants focus on the correct picture first, then looked increasingly more at the wrong - more salient - stimulus, and then went back to the correct stimulus. 

```{r results = FALSE}
dat_attention_plot <- describeBy(dat_attention_long$mean_accuracy,
                                 list(dat_attention_long$Condition,
                                      dat_attention_long$frame),
                                 mat = TRUE, digits = 2)
names(dat_attention_plot)[names(dat_attention_plot) == 'group1'] = 'condition'
names(dat_attention_plot)[names(dat_attention_plot) == 'group2'] = 'frame'
```

<b> Figure 1 </b>
```{r, fig.width = 9, fig.height = 4}
ggplot(dat_attention_plot, aes(x = sort(frame, decrease = FALSE), y = mean, group = condition, linetype = condition)) +
  geom_line(stat = "identity") +
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se, fill = "Mean +/- 1SE"), alpha = 0.2) +
  scale_fill_manual("", values = "grey12") +
  ylab("Looks to Target") +
  xlab("") +
  apatheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c("frame_087","frame_097", "frame_148"),
                   labels = c("0 ms", "300 ms", "2000 ms")) +
  geom_vline(xintercept = 97, linetype = "dashed") +
  geom_vline(xintercept = 148, linetype = "dashed")
```

# Time series analysis

```{r}
dat_attention_long$frame <- as.character(dat_attention_long$frame)
dat_attention_long$frame.n <- substr(dat_attention_long$frame, 7, nchar(dat_attention_long$frame))
dat_attention_long$frame.n <- as.numeric(dat_attention_long$frame.n)
```

```{r}
dat_attention_long$Tr_Num <- as.factor(dat_attention_long$Tr_Num)
dat_attention_long$mean_accuracy.c <- scale(dat_attention_long$mean_accuracy, center = TRUE, scale = FALSE)
model_time_sequence <- lmer(mean_accuracy.c ~ Condition * frame.n + ( 1 | Child_ID) + ( 1 | Tr_Num),
                            data = dat_attention_long, REML = FALSE)
summary(model_time_sequence)
```

