---
title: "Visual attention to neutral vs competing vs reinforcing stimuli among children with ASD"
output: html_notebook
---
<b>Author: Dominik Neumann</b><br>
Michigan State University

```{r global_options, include = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Setup

Please select "dat_visualattention.csv" for dat_attention data table and "dat_master.csv" for dat_master data table in order to run this script properly. <br>
Missing values were coded as -999. We coded “-“ values as -888, indicating participants looked at the center of the screen and thus at neither distracter nor target. We coded “.” values as -777, indicating participants looking outside of the screen.

```{r results = FALSE}
### OPTIONS
options(scipen = 3)

### LOADING PACKAGES
# General libraries
library(tidyverse)
library(psych)
library(plyr)

# Multilevel modeling
library(lme4)
library(nlme)
library(ez)
library(multcomp)
library(TTR)
library(lmerTest)
library(powerlmm)
library(lsmeans)
library(car)

# ANOVA
library(knitr)
library(sjstats)

### FUNCTIONS
# APA style charts
apatheme <- theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(),
        text = element_text(family = 'Times'))

# Function that counts the number if NA's in a given row:
countNA <- function(df)
  apply(df, MARGIN = 1, FUN = function(x) length(x[is.na(x)]))

### DATA
# Loading gaze data
dat_attention <- read.csv(file.choose())
dat_master <- read.csv(file.choose())

# Recoding '-999', '-888', and '-777' to missing values 'NA'
dat_attention[dat_attention == -999 | dat_attention == -888 | dat_attention == -777] <- NA
```

There are 58 participants in this study that have been diagnosed with Autism Syndrom Disorder (ASD). All conditions were administered within participants, with ten trials per condition.

```{r results = FALSE}
# We were only interested in participants with with Autism Syndrom Disorder
dat_attention$Child_ID <- factor(dat_attention$Child_ID)
dat_attention <-dat_attention %>%
  dplyr::filter(Child_Type == "ASD")
nrow(count(unique(dat_attention$Child_ID)))
```

# Data cleaning and transforming

```{r}
# Drop columns before and after the analysis window (3200 = frame 97, 4900 = frame 148)
dat_attention_analysiswindow <- subset(dat_attention, select = -c(frame_001:frame_096, frame_149:frame_200))
```

## Missing values and underperforming trials

Our critical window is from 300 to 2000 ms after the onset of the target word. Trials on which children were not fixating either object for more than 50% of the critical window were excluded. These trials do not have adequate data. In sum, there were 1,313 trials. Of these, 397 (~30%) had more than 50% missing data. We further excluded 67 trials because participants with less than four valid trials were excluded from further analysis. After exclusion we therefore used 849 trials.

```{r results = FALSE}
### TRIAL-WISE DELETION
# Count number of rows before deleting trials with more than 50% missing values
nrow(dat_attention_analysiswindow)
# Apply function countNA (see in "Functions") to critical dataset and exclude all rows with more than (number of variables divided by two) NAs between frame_097 and frame_148
dat_attention_analysiswindowNA <- dat_attention_analysiswindow[countNA(subset(dat_attention_analysiswindow,
                                                                              select = frame_097:frame_148)) < (ncol(subset(dat_attention_analysiswindow,
                                                                                                                            select = frame_097:frame_148))/2),]
# Count number of rows after deleting trials with more than 50% missing values
nrow(dat_attention_analysiswindowNA)
nrow(dat_attention_analysiswindow) - nrow(dat_attention_analysiswindowNA)
(nrow(dat_attention_analysiswindow) - nrow(dat_attention_analysiswindowNA)) / (nrow(dat_attention_analysiswindow))
```

<b>Table 1</b> Number of trials after excluding trials with more than 50% missingness
```{r}
Process <- c("Before cleaning", "Number of excluded trials", "After cleaning")
n <- c(1313, 397, 916)
number_of_observations <- cbind(Process, n)
as.data.frame(number_of_observations)
```

```{r results = FALSE}
### PARTICIPANT-WISE DELETION
# Counting number of trials within conditions for each participant before exclusion
kidsxconditiontable <- count(dat_attention_analysiswindowNA, c("Child_ID", "Condition"))
nrow(kidsxconditiontable)

# Excluding participant who were contributing to little
id_for_exclusion <- c("8016", "8022", "8047", "8049", "8057", "8058", "8059", "8062", "8063", "8072", "8076", "8077")
dat_attention_analysiswindowNA <- dat_attention_analysiswindowNA[ ! dat_attention_analysiswindowNA$Child_ID %in% id_for_exclusion, ]

# Counting number of trials within conditions for each participant after exclusion
kidsxconditiontable1 <- count(dat_attention_analysiswindowNA, c("Child_ID", "Condition"))
nrow(kidsxconditiontable1)
916 - nrow(dat_attention_analysiswindowNA)
nrow(dat_attention_analysiswindowNA)
```

## Overview of missing data
After excluding participants with more than 50% missingness, participants had on average 4.83 (SD = 6.13) missing values for the eye tracking data (participant did not look at either of the pictures). There were significantly less missing values in the reinforcing condition as compared to the competing condition (&Delta; = -1.51, <i>CI</i> [-2.80, -0.23], <i>p</i> = 0.02) and as compared to the neutral condition (&Delta; = -1.43, <i>CI</i> [-2.74, -0.19], <i>p</i> = 0.02). This might be due to the fact that there were less observations in the reinforcing condition than in the other two condition, because of fewer number of trials (N<sub>competing</sub> = 316, N<sub>neutral</sub> = 329, N<sub>reinforcing</sub> = 204). That is, there were 6 trials in the reinforcing condition, but 10 trials in the competing and neutral condition respectively.<br>

<b>Table 2</b> Missing values per trial (overall)
```{r results = FALSE}
dat_attention_analysiswindow_descriptive <- dat_attention_analysiswindowNA %>%
  dplyr::select("Condition", starts_with("frame_"))
dat_attention_analysiswindow_descriptive$NAcount <- countNA(dat_attention_analysiswindow_descriptive)
describe(dat_attention_analysiswindow_descriptive$NAcount)
```

<b>Table 3</b> Missing values (stratified by condition)
```{r}
# Summary
descriptive_missing <- describeBy(dat_attention_analysiswindow_descriptive$NAcount,
                                  dat_attention_analysiswindow_descriptive$Condition,
                                  mat = TRUE, digits = 2)
descriptive_missing
anova_missingvalues <- aov(NAcount ~ Condition, data = dat_attention_analysiswindow_descriptive)
TukeyHSD(anova_missingvalues)
```

## Overview of trials per participant

Participants completed a mean of 7.1 (SD = 1.95) trials per condition.<br>

<b>Table 4</b> Summary trials per participants
```{r results = FALSE}
dat_attention_analysiswindowNA$Child_Condition <- interaction(dat_attention_analysiswindowNA$Child_ID,
                                                              dat_attention_analysiswindowNA$Condition)
# Creating variable containing counts of trials retained per participant
dat_TRcounts <- ddply(dat_attention_analysiswindowNA,
                      ~Child_Condition,
                      summarise,
                      CountsTR = length(unique(Tr_Num)))

dat_attention_analysiswindowNA <- merge(dat_attention_analysiswindowNA,
                                        dat_TRcounts,
                                        by = "Child_Condition")

describe(dat_attention_analysiswindowNA$CountsTR)
```

<b>Table 5</b> Completed trials per participant (strafied by condition)
```{r}
# Summary
descriptive_trial <- describeBy(dat_attention_analysiswindowNA$CountsTR,
                                dat_attention_analysiswindowNA$Condition,
                                mat = TRUE, digits = 2)
descriptive_trial
```

## Transformation to long data

```{r results = FALSE}
# Aliasing data
dat_analysiswindow <- dat_attention_analysiswindowNA
```

In order to analyse data appropriately, data was transformed to a series of long datasets. First, a dataset for the analysis window only (frame 97 to frame 148) was created. Then, a dataset for all time points (frame 1 to frame 200) was created. Additionally, a dataset was generated that included the time before the analysis window as well as after the analysis window.<br>

```{r results = FALSE}
### TRANSFORMING ANALYSIS WINDOW ONLY
dat_analysiswindow$Child_ID <- factor(dat_analysiswindow$Child_ID)
dat_analysiswindow_long <- gather(dat_analysiswindow,
                                  frame, accuracy,
                                  frame_097:frame_148,
                                  factor_key = TRUE)

### OVERALL DATA
# Subsetting full data based on prior cleaning procedures
dat_attention_full <- subset(dat_attention, Child_ID %in% dat_analysiswindow$Child_ID)

# Subsetting for the pre-analysis window
dat_attention_pre <- subset(dat_attention_full, select = -c(frame_097:frame_200))
dat_attention_pre$prepost <- "pre-analysis"
dat_attention_pre$Child_ID <- factor(dat_attention_pre$Child_ID)
dat_attention_pre_long <- gather(dat_attention_pre,
                                 frame, accuracy,
                                 frame_001:frame_096,
                                 factor_key = TRUE)

# Subsetting for the analysis window
dat_attention_analysis <- subset(dat_attention_full, select = -c((frame_001:frame_096), (frame_149:frame_200)))
dat_attention_analysis$prepost <- "analysis"
dat_attention_analysis$Child_ID <- factor(dat_attention_analysis$Child_ID)
dat_attention_analysis_long <- gather(dat_attention_analysis,
                                      frame, accuracy,
                                      frame_097:frame_148,
                                      factor_key = TRUE)

# Subsetting for the post-analysis window
dat_attention_post <- subset(dat_attention_full, select = -c((frame_001:frame_148)))
dat_attention_post$prepost <- "post-analysis"
dat_attention_post$Child_ID <- factor(dat_attention_post$Child_ID)
dat_attention_post_long <- gather(dat_attention_post,
                                  frame, accuracy,
                                  frame_149:frame_200,
                                  factor_key = TRUE)

# Combining
dat_attention_long <- rbind(dat_attention_pre_long,
                            dat_attention_analysis_long,
                            dat_attention_post_long)
```

# Measuring Accuracy

"ACCURACY" is measured as '0', '1', or 'NA'.<br>
1 = participant focuses on target at specific frame<br>
0 = participant focuses on the distracter at specific frame<br>
NA = participant focuses either on a point outside the screen (wall, machines, roof, etc.) or on the center of the screen (neither picture) at specific frame<br>
"MEAN ACCURACY" is measured as the mean of accuracy observations across individuals for each frame within a given conditions.

```{r results = FALSE}
# For the analysis window dataframe
dat_analysiswindow_long$accuracy <- as.numeric(as.character(dat_analysiswindow_long$accuracy))
dat_analysiswindow_long <- dat_analysiswindow_long %>%
  dplyr::group_by(Child_ID, frame, Condition) %>%
  dplyr::mutate(mean_accuracy = mean(accuracy, na.rm = TRUE))
# For the overall dataframe
dat_attention_long$accuracy <- as.numeric(as.character(dat_attention_long$accuracy))
dat_attention_long <- dat_attention_long %>%
  dplyr::group_by(Child_ID, frame, Condition) %>%
  dplyr::mutate(mean_accuracy = mean(accuracy, na.rm = TRUE))
```

# Plotting Mean Accuracy by Condition

Figure 1 depicts development of mean accuracy. That is, the y-axis describes mean accuracy for all participants over all trials and the x-axis describes the time from beginning to end of the experiment. Mean accuracy measures are stratified by condition. We can see that children generally paid more attention to the correct stimulus in the "reinforcing condition". Children further paid the least attention to the correct stimulus in the competing condition. However, an interesting orienting pattern emerged in the beginning, where participants focus on the correct picture first, then looked increasingly more at the wrong - more salient - stimulus, and then went back to the correct stimulus. 

```{r results = FALSE}
dat_attention_kidxcondition <- describeBy(dat_attention_long$mean_accuracy,
                                          list(dat_attention_long$Condition,
                                               dat_attention_long$frame),
                                          mat = TRUE, digits = 2)
names(dat_attention_kidxcondition)[names(dat_attention_kidxcondition) == 'group1'] = 'condition'
names(dat_attention_kidxcondition)[names(dat_attention_kidxcondition) == 'group2'] = 'frame'
```

<b> Figure 1 </b>
```{r, fig.width = 9, fig.height = 4}
ggplot(dat_attention_kidxcondition,
       aes(x = sort(frame, decrease = FALSE), y = mean, group = condition, linetype = condition)) +
  geom_line(stat = "identity") +
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se, fill = "Mean +/- 1SE"), alpha = 0.2) +
  scale_fill_manual("", values = "grey12") +
  ylab("Looks to Target") +
  xlab("") +
  apatheme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c("frame_087","frame_097", "frame_148"),
                   labels = c("0 ms", "300 ms", "2000 ms")) +
  geom_vline(xintercept = 97, linetype = "dashed") +
  geom_vline(xintercept = 148, linetype = "dashed")
```

